Introduction: 

Introducing wireless communications in the network world showed us the evolution of how we came from the demands of modern digital applications which have increased. The generation past us of wireless technology introduced significant improvements, from mobile broadband in 4G networks to massive connectivity and low-latency services in 5G networks. However, if we talk about the future now, there is growth being done with transportation, and with health systems requiring and needing new sets of networks which are capable of able to manage the extreme levels of connectivity, computation, and automation.

The 6G of network evolution is about to spiral rapidly as time goes by. The use of artificial intelligence will be even more advanced because of the use of 6G aiming to integrate AI as a fundamental component of network operation. The overall concept of it is the AI native networking which enables the network to optimize resources, analyze conditions and make even more intelligent decisions with minimal human intervention. The research will show the ideas of combining machine learning, distributed intelligence, and advanced communication technologies to create highly adaptive wireless environments for the future. The AI integration will give us many advantages for the future of wireless systems with advantages like having autonomous control, resource management and having better network optimization to make it easier to load. The tasks will be easier to do because it will be assisted by artificial intelligence which will help in network automation, channel estimation and more. The power of these capabilities are very important for the future of the network to be able to support billions. Supporting the billions by being able to generate large amounts of data and requiring real-time decision-making for all devices in real time. 

With all of these advantages coming and playing through challenges are also up the arisen in which we need to also deal with. Challenges that can rise can range from privacy, security, and consumer trust. The challenges can come from the concerns from the data ownership, transparency, and responsible AI usage. The continuous collection of the data would help us determine the essentials of what challenges of what will come. Having to understand the overall foundation of  AI-native 6G networks is important to able to see what solutions we can figure in the future for the challenges that are associated with future wireless services.

Background: AI-Native 6G Networks: 

	The development of wireless communication has progressed from different generations with also introducing new capabilities and services as time flies. With new capabilities coming, the 5G network has only been introduced with communication, ultra-reliable low-latency communication, the 6G network is about to be even better. With the capabilities of being able to provide even more advanced capabilities through intelligent and independent operations. With even more support on high data rates, ultra-low latency communication, massive device connectivity, and integration between communication and sensing capabilities. The capability of supporting the applications, can introduce the concept that wireless signals are used not only for transmitting information but also for sensing the surrounding environment. A positive impact where it can provide which will be monitored to improve in the future. 

	AI networking would play a big shift in the wireless network world as we see. Artificial intelligence has the tool to be an external optimization tool which can be integrated into the network architecture. The network can continuously learn from collected data, adapt to changing conditions, and automatically optimize performance. AI-native 6G networks can be able to use algorithms that can be able to improve any network functions, including resource allocation, traffic management, energy efficiency, and network reliability. An example we can state is how AI can be able to predicate any changes that can occur between the network demand and automatically adjust available resources to maintain performance. 

	AI-native 6G networks rely on several AI technologies to improve performance and automation. Machine learning helps analyze network data, predict traffic, optimize resources, and enhance decision-making. Deep learning supports complex wireless tasks such as signal processing and channel estimation. Reinforcement learning allows networks to learn from their environment and make better decisions over time. Federated learning improves privacy by enabling devices to train AI models locally while sharing only updates instead of raw data. 


	Although AI-native 6G networks offer many benefits, challenges still exist. Large-scale data use raises privacy and user control concerns, while AI decision-making can create issues with transparency, fairness, and accountability. Protecting AI models from attacks and ensuring reliable performance are also major research challenges. Addressing these concerns is essential for building secure and trustworthy AI-driven 6G networks. 
